CrypTool Script \\ Mathematics and Cryptography

### 2001-12-20 script-gb/script-gb.tex
no message

\title{ CrypTool Version 1.3 \\ Tutorial }

\author{
(c) Deutsche Bank AG 1998-2001 \\
Frankfurt am Main
}

@@

You will receive information about the principles of symmetrical and
asymmetrical encryption. A large section of the tutorial is dedicated to the
fascinating topic of prime numbers. XXX You will also obtain an overview of the
mathematical ideas behind modern cryptography.XXX

@@

The authors Bernhard Esslinger, Bartol Filipovic, Henrik Koy and Roger Oyono
would like to take this opportunity to thank their colleagues in the company and
at the universities of Frankfurt, Gießen und Karlsruhe. They particularly thank
Dr. Peer Wichmann from the Karlsruhe computer science research centre
(Forschungszentrum Informatik, FZI) for his down-to-earth support.

@@

A further aim was to enable users to understand the cryptographic procedures
implemented in the Deutsche Bank. XXX In this way, using CrypTool as a reliable
reference implementation of the various encryption procedures, you can test the
encryption implemented in other programs. XXX \par \vskip + 3pt

@@

\newpage
\section{Prime numbers}
(Bernhard Esslinger, May 1999, Update Nov. 2000)

@@

Due to the fact that the exponents of the Mersenne numbers do not use all
natural numbers, but only the primes, the {\em experimental space} is limited
considerably. The currently known Mersenne prime numbers have the exponents $$
\begin{array}{c}
2, {} 3, {} 5, {} 7, {} 13, {} 17, {} 19, {} 31, {} 61, {} 89, {} 107, {} 127,
{} 521, {} 607, {} 1.279, {} 2.203, {} 2.281, {} 3.217, {} 4.253, {} 4.423, \\
9.689, {} 9.941, {} 11.213, {} 19.937, {} 21.701, {} 23.207, {} 44.497, {}
86.243, {} 110.503, {} 132.049, {} 216.091,\\
 756.839, {} 859.433, {} 1.257.787, {} 1.398.269, {} 2.976.221, {} 3.021.377, {}
6.972.593.
\end{array}
$$
Thus $38$ Mersenne prime numbers are currently known. The $19$th number with the
exponent $4,253$ was the first with at least $1,000$ digits in  decimal system
(the mathematician Samual \index{Yates Samual} Yates coined the expression {\em
titanic} \index{Prime numbers!titanic} prime for this; it was discovered by
Hurwitz in 1961); the $27$th number with the exponent $44,497$ was the first
with at least $10,000$ digits in the decimal system (Yates coined the expression
\index{Prime numbers!gigantic}  {\em gigantic} prime for this. These names are
now long outdated). The 37th number was found in January 1998 and has 909,526
digits in the decimal system, which corresponds to 33 pages in the newspaper!

@@

{\href{http://reality.sgi.com/chongo/prime/prime_press.html}{\tt http://reality.sgi.com/chongo/prime/prime\_press.html}}

\begin{quote}
(The supercomputer manufacturer SGI Cray Research not only employed brilliant
mathematicians but also used the prime number tests as benchmarks for their
machines.)
\end{quote}
{\href{http://www.utm.edu/}{\tt http://www.utm.edu/}}
\begin{quote}
(At the university of Tennessee you will find extensive research results about
prime numbers.)
\end{quote}
The largest prime number currently known (November 2000) is the Mersenne number
$$ 2^{6,972,593} - 1. $$
It was discovered in June 1999 and has $2,098,960$ digits in the decimal system
(that corresponds to around 77 pages in the newspaper). Like the three Mersenne
numbers listed above, this number was also discovered in the GIMPS project
(Great Internet Mersenne Prime Search). Since 1995, dozens of experts and
thousands of amateurs have been working on this project to find such numbers
using distributed computer programs.

@@

{\href{http://www.eff.org/coopawards/prime-release1.html}{\tt http://www.eff.org/coopawards/prime-release1.html}},

%{\href{http://www.octocad.demon.co.uk/mersenne/prime.htm }{\tt http://www.octocad.demon.co.uk/mersenne/prime.htm}}.

@@

This is the \index{Prime number theorem} \textbf{prime number theorem}. It was
put forward by Legendre (1752-1833) and Gauß (1777-1855) but not proved until
over 100 years later.

@@

\begin{itemize}
  \item For each number $n$ in ${\bf N}$ there are $n$ consecutive natural
numbers that are not primes.
    A proof of this can be found in \cite[p. 79]{Padberg1996}.
  \item The Hungarian mathematician Paul Erdös (1913-1996) proved:
    Between each random number not equal to $1$ and its double, there is at
least one prime. (He was not the first to prove this theorem, but proved it in a
simpler manner than those before him).
  \item There is a real number a such that the function $f: {\bf N} \rightarrow
{\mathbb Z}$ where $n \mapsto a^{3^n}$     only delivers primes for all $n$ (see
\cite[p. 82]{Padberg1996}).
    Unfortunately, problems arise when we try to determine $a$ (see below).
\end{itemize}

@@

\newpage\begin{thebibliography}{99999}
\addcontentsline{toc}{subsection}{Bibliography}
\bibitem[Bartholome1996]{Bartholome1996}  \index{Bartholome 1996}     A.
Bartholomé, J. Rung, H. Kern, \\     Zahlentheorie für Einsteiger, Vieweg 1995,
2. Auflage 1996.

@@

\bibitem[Klee1997]{Klee1997} \index{Klee 1997}     V. Klee, S. Wagon, \\
Ungelöste Probleme in der Zahlentheorie und der     Geometrie der Ebene,
Birkhäuser Verlag, 1997.

@@

\bibitem[Schwenk1996]{Schwenk1996} \index{Schwenk 1996}     J. Schwenk \\
Conditional Access. In: taschenbuch der telekom praxis 1996, Hrgb. B. Seiler,
Verlag Schiele und Schön, Berlin.

\bibitem[Tietze1973]{Tietze1973} \index{Tietze 1973}     H. Tietze, \\
Gelöste und ungelöste mathematische Probleme, Verlag C.H. Beck,     1959, 6.
Auflage 1973.

@@

\newpage
\section{Elliptic Curves}
\index{Elliptic curves}
\hypertarget{ellcurve}{}
\subsection{Elliptic curves --- history}

@@

Because the term {\em group} is used differently in everyday language than in
mathematics, we will, for reasons of completeness, begin by introducing the
essential statement of the formal definition of a group:
\begin{itemize}
   \item A group is a non-empty set $G$ and an operation $+.$ The set $G$ is
closed under the operation $+.$ Regardless of which two elements $a, b$ from $G$
are taken, performing the operation on them gives an element in $G$ (i.e.
$a+b=c$, and $c$ lies in $G$).
   \item For all elements $a, b$ and $c$ in $G$: $(a+b)+c = a+(b+c)$.
   \item There exists an element $e$ in $G$ that behaves neutrally with respect
to the operation $+$. That means that for all a in the set G: $a+e = e+a = a.$
   \item For each element $a$ in $G$ there exists a so-called inverse element $-
a$ ($-a$ also lies in $G$) such that: $a+(-a) = (-a)+a = e$.
\end{itemize}
If also $a+b = b+a$ for all $a, b$ in $G$ then we call the group an {\em
Abelian} group. An operation denoted as $+$ indicates an {\em additive} group;
if the operation is denoted as $\cdot$, we speak of a {\em multiplicative}
group.

@@

An elliptic curve over ${\mathbb Z}_p$ is defined by an equation of the
following form: $$ y^2 \; ({\rm mod} \; p) = x^3 + ax + b ({\rm mod} \; p) $$
(thus: equality in the field ${\mathbb Z}_p$), where $a, b$ are in ${\mathbb
Z}_p$ and $4a^3 + 27b^2$ mod $p$ is not equal to zero. For fixed chosen numbers
$a$ and $b$ in ${\mathbb Z}_p$, this equation has the pair of solutions $$ {\bf
E} = \left\{(x,y) \left| \begin{array}{c} x {\rm ~and~} y {\rm ~are~in~}
{\mathbb Z}_p {\rm ~and~} \\ y^2  \equiv x^3 + ax + b \; ({\rm mod~} p) {\rm
~and~} \\ 4a^3 + 27b^2 \not\equiv 0 \; ({\rm mod} p)\end{array} \right.
\right\}, $$ i.e. the set ${\bf E}$ consists of all pairs $x$ and $y$ that are a
solution (in ${\mathbb Z}_p$) of the above equation. It must be noted that the
numbers $a, b$ and $p$ determine which pairs $(x,y)$ lie in the set ${\bf E}$.
This means that $a, b$ and $p$ specify this set. The elements $(x,y)$ in ${\bf
E}$ are called points on the elliptic curve. In addition, ${\bf E}$ has one more
element $O$ (the so-called point in infinity). The set ${\bf E}$ is usually
called an elliptic curve. We can now define an operation (also denoted as $+$,
although it is not the standard/usual addition of real numbers) on two elements
in ${\bf E}$ such that the operation delivers an element that also lies in ${\bf
E}$. The set ${\bf E}$ is therefore closed under the operation $+$. We can show
that ${\bf E}$ is a group. The neutral element of the group ${\bf E}$ is the
point in infinity $O$. Thus, for every two points $(x_1,y_1)$ and $(x_2,y_2)$ on
the elliptic curve ${\bf E}$, there exists a point $(x_3,y_3)$ on ${\bf E}$ such
that the operation $+$ complies with the following: $(x_1,y_1) + (x_2,y_2) =
(x_3,y_3)$. Under certain circumstances, these points may also be equal to the
point in infinity. Thus, when we speak of a point $P$ on an elliptic curve ${\bf
E}$, we mean that $P = (x,y)$ and $(x,y)$ lies in the set ${\bf E}$. Any two
points on an elliptic curve specified by $a, b$ and $p$ can therefore be added
and the result is a point that also lies on the same elliptic curve. We must
note that ${\bf E}$ can have the following meanings:
\begin{itemize}
   \item the set ${\bf E}$ of solutions pairs $(x,y)$ for an equation including the point $O$
   \item the group ${\bf E}$ (with the operation ''addition of $(x_1,y_1)$ and $(x_2,y_2)$'')
   \item the elliptic curve ${\bf E}$ (which is actually the same as the group ${\bf E}$)
 \end{itemize}

@@

For cryptography, the important fact is that, for very large
numbers, it appears to be extremely difficult to use a given point
$Q$ on an elliptic curve to determine which two points have to be
added to obtain $Q$. For large numbers $a, b$ and $p$ ($p$, for
example, has a length of more than $160$ bits), the computer can
easily add the point $P$  $m$ times after another, i.e. to
determine the point $P + P + \cdots + P = Q$ in an incredibly
short space of time (in a few fractions of a second) ($m$ summands
$P$). Rather than $P + P + \cdots + P = Q$ ($m$ summands $P$) we
also write $mP = Q$. If we have a point $P$ and a point $Q$, which
both lie on the same elliptic curve, no procedure is known that
enables us --- within an acceptable space of time --- to determine
the number $m$ (assuming it actually exists) for which $mP = Q$.
This is referred to as the ''elliptic curve discrete logarithm
problem'' (abbreviated to ECDLP).

We must note that not all elliptic curves are equally secure. This
means that we must choose the parameters $a$ and $b$ carefully
when defining a curve. For certain classes of elliptic curves, it
is possible to solve the ECDLP easier than in the general case.
Cryptographically unsuitable elliptic curves are called abnormal
curves (these are curves over ${\mathbb Z}_p$ for which the set
${\bf E}$ has precisely $p$ elements) and the supersingular curves
(the curves for which we can reduce the calculation of the ECDLP
to calculating the ''standard'' discrete logarithm in other finite
fields, i.e. simplify the calculation). There are therefore
cryptographically good and bad curves. However, for given
parameters $a$ and $b$ we can, with some difficulty, establish
whether or not the resulting elliptic curve is useful for
cryptographic purposes. The curves used in cryptography are
usually provided by experts. They ensure that the elliptic curves
they classify as secure satisfy the current security requirements.
For secure curves, the parameter $p$ determines how long it takes
to solve the ECLDP on this curve. The larger the parameter $p$,
the longer it takes to solve the problem. Experts recommend a bit
length of over $200$ bits for the parameter $p$. This makes it
clear why elliptic curves are so interesting for cryptography.
Because the parameter $p$ also determines the time required to
perform the signature/encryption procedure when using elliptic
curves in cryptography. The time taken to generate a pair of keys
also depends on $p$. Thus, small values (few bits) are desirable
here (in order to minimise the run times for the procedures);
however, the required security must still be maintained. For
example, with a length of $200$ bits for $p$, a {\em good}
elliptic curve is just as secure as an \index{RSA!module} RSA
module of over $1024$ bits in length (at least according to the
current state of research). The reason for this is that the
quickest algorithms for solving the {\em elliptic curve discrete
logarithm} problem have an exponential run time --- unlike the
subexponential run times that the best factorisation algorithms
currently have (number sieve, quadratic sieve or factorisation
with elliptic curves). Therefore, the parameters for cryptographic
procedures based on the problem of {\em factorising whole numbers}
must be greater than the parameters for cryptographic procedures
based on the ECDL problem.

@@

The {\em elliptic curve discrete logarithm problem} (ECLDP) forms the basis for
elliptic-curve cryptography. Various signature procedures are based on this.
What they have in common is how they use the public parameters and how these
parameters are used to generate secret and public keys:
\begin{itemize}
    \item The parameters of the elliptic curve E, i.e. a prime number $p$ that
determines over which field ${\mathbb Z}_p$ the elliptic curve ${\bf E}$ is
defined, as well as the two numbers $a$ and $b$ in ${\mathbb Z}_p$.
    \item A point $G=(x,y)$ that lies on the elliptic curve ${\bf E}$.
    \item A prime number $r < p$ for which $rG=O$ (i.e. $G$ added $r$ times
gives the neutral element of the group ${\bf E}$) and $r$ is a factor of $\#{\bf
E}$ (where $\# {\bf E}$ is the number of elements in the set ${\bf E}$). The
point $G$ therefore has the order $r$ and is the generator of a cyclic subgroup
of ${\bf E}$ with the order $r$.
    \item The number $k = \#{\bf E}/r$ ($k$ is called the {\em cofactor}).
\end{itemize}
The parameters $a, b, p, G, r$ and $k$ listed above are called
domain parameters. They determine on which elliptic curve ${\bf
E}$ and in which cyclic subgroup of ${\bf E}$ a signature
procedure has been ''used''.

>>>>

\title{ CrypTool Script \\ Mathematics and Cryptography }

\author{
(c) The authors, 1998-2001 \\
Frankfurt am Main
}

@@

You will receive information about the principles of symmetrical and
asymmetrical encryption. A large section of the tutorial is dedicated to the
fascinating topic of prime numbers. You will also obtain an overview of the
mathematical ideas behind modern cryptography.

@@

The authors Bernhard Esslinger, Bartol Filipovic, Henrik Koy and Roger Oyono
would like to take this opportunity to thank their colleagues in the company and
at the universities of Frankfurt, Gie\"sen und Karlsruhe. They particularly thank
Dr. Peer Wichmann from the Karlsruhe computer science research centre
(Forschungszentrum Informatik, FZI) for his down-to-earth support.

@@

A further aim was to enable users to understand the cryptographic procedures
implemented in the Deutsche Bank. In this way, using CrypTool as a reliable
reference implementation of the various encryption procedures, you can test the
encryption implemented in other programs. \par \vskip + 3pt

@@

\newpage
\section{Prime numbers}\hypertarget{Kapitel_2}{}
(Bernhard Esslinger, besslinger@web.de, May 1999, Update Nov. 2000, December 2001)

@@

Due to the fact that the exponents of the Mersenne numbers do not use all
natural numbers, but only the primes, the {\em experimental space} is limited
considerably. The currently known Mersenne prime numbers have the exponents $$
\begin{array}{c}
2, {} 3, {} 5, {} 7, {} 13, {} 17, {} 19, {} 31, {} 61, {} 89, {} 107, {} 127,
{} 521, {} 607, {} 1.279, {} 2.203, {} 2.281, {} 3.217, {} 4.253, {} 4.423, \\
9.689, {} 9.941, {} 11.213, {} 19.937, {} 21.701, {} 23.207, {} 44.497, {}
86.243, {} 110.503, {} 132.049, {} 216.091,\\
 756.839, {} 859.433, {} 1.257.787, {} 1.398.269, {} 2.976.221, {} 3.021.377, {}
6.972.593, ~13.466.917.
\end{array}
$$
Thus $39$ Mersenne prime numbers are currently known.

The $19$th number with the exponent $4,253$ was the first with at least $1,000$ digits in  decimal system
(the mathematician Samual \index{Yates Samual} Yates coined the expression {\em
titanic} \index{Prime numbers!titanic} prime for this; it was discovered by
Hurwitz in 1961); the $27$th number with the exponent $44,497$ was the first
with at least $10,000$ digits in the decimal system (Yates coined the expression
\index{Prime numbers!gigantic}  {\em gigantic} prime for this. These names are
now long outdated).

The 37th number was found in January 1998 and has 909,526
digits in the decimal system, which corresponds to 33 pages in the newspaper!

@@

{\href{http://reality.sgi.com/chongo/prime/prime_press.html}{\tt http://reality.sgi.com/chongo/prime/prime\_press.html}}
\begin{quote}
(The supercomputer manufacturer SGI Cray Research not only employed brilliant
mathematicians but also used the prime number tests as benchmarks for their
machines.)
\end{quote}
{\href{http://www.utm.edu/}{\tt http://www.utm.edu/}}
\begin{quote}
(At the university of Tennessee you will find extensive research results about
prime numbers.)
\end{quote}
\subsubsection{M-38 -- June 1999}\index{Mersenne Marine!M-38}
The 38th Mersenne prime 
$$ 2^{6,972,593} - 1. $$
was discovered in June 1999 and has $2,098,960$ digits in the decimal system
(that corresponds to around 77 pages in the newspaper).

\subsubsection{M-39 -- December 6, 2001} \index{Mersenne Marine!M-39}

The 39th Mersenne prime, called M-39, ($=2^{13.466.917}-1$), was found at December 6, 2001 – more exactly,
he verification of this number, found at November 14, 2001 by the Canadian student Michael Cameron,
was successfully finished. This number has about 4 million decimal digits (exactly 4,053,946 digits).
Trying only to print this number (924947738006701322247758 ... 1130073855470256259071) in the Financial Times
would need around 200 pages.

\subsubsection{GIMPS}\index{GIMP}

Discovering the 39th Mersenne prime the GIMPS project (Great Internet Mersenne Prime Search) founded in 1996
already discovered for the 5th time the greatest Mersenne number to be proven being prime.
Now more than 130,000 volunteers, amateurs and experts, are working for the GIMPS project. They connect their
computers into the so called „primenet“, organized by the company entropia, to find such numbers using distributed
computer programs.

\subsubsection{EFF}\index{EFF}

@@

{\href{http://www.eff.org/coopawards/prime-release1.html}{\tt http://www.eff.org/coopawards/prime-release1.html}}.

The discoverer of M-38 got for discovering the first prime with more than 1 million decimal digits 50.000 USD 
from the EFF. The next price from EFF with 100.000 USD is for a proven prime with more than 10 million decimal digits.
%{\href{http://www.octocad.demon.co.uk/mersenne/prime.htm }{\tt http://www.octocad.demon.co.uk/mersenne/prime.htm}}.

@@

This is the \index{Prime number theorem} \textbf{prime number theorem}. It was
put forward by Legendre (1752-1833) and Gau\"s (1777-1855) but not proved until
over 100 years later.

@@

\begin{itemize}
  \item For each number $n$ in ${\bf N}$ there are $n$ consecutive natural
numbers that are not primes.
    A proof of this can be found in \cite[p. 79]{Padberg1996}.
  \item The Hungarian mathematician Paul Erd\"os (1913-1996) proved:
    Between each random number not equal to $1$ and its double, there is at
least one prime. (He was not the first to prove this theorem, but proved it in a
simpler manner than those before him).
  \item There is a real number a such that the function $f: {\bf N} \rightarrow
{\mathbb Z}$ where $n \mapsto a^{3^n}$     only delivers primes for all $n$ (see
\cite[p. 82]{Padberg1996}).
    Unfortunately, problems arise when we try to determine $a$ (see below).
\end{itemize}

@@

\newpage\begin{thebibliography}{99999}
\addcontentsline{toc}{subsection}{Bibliography}
\bibitem[Bartholome1996]{Bartholome1996}  \index{Bartholome 1996}     A.
Bartholomé, J. Rung, H. Kern, \\     Zahlentheorie f\"ur Einsteiger, Vieweg 1995,
2. Auflage 1996.

@@

\bibitem[Klee1997]{Klee1997} \index{Klee 1997}     V. Klee, S. Wagon, \\
Ungel\"oste Probleme in der Zahlentheorie und der     Geometrie der Ebene,
Birkh\"auser Verlag, 1997.

@@

\bibitem[Schwenk1996]{Schwenk1996} \index{Schwenk 1996}     J. Schwenk \\
Conditional Access. In: taschenbuch der telekom praxis 1996, Hrgb. B. Seiler,
Verlag Schiele und Sch\"on, Berlin.

\bibitem[Tietze1973]{Tietze1973} \index{Tietze 1973}     H. Tietze, \\
Gel\"oste und ungel\"oste mathematische Probleme, Verlag C.H. Beck,     1959, 6.
Auflage 1973.

@@

\newpage
\section{Elliptic Curves}
\index{Elliptic curves}
\hypertarget{ellcurve}{}
(B. Filipovics / B. Esslinger, April 2000)

\subsection{Elliptic curves --- history}

@@

Because the term {\em group} is used differently in everyday language than in
mathematics, we will, for reasons of completeness, begin by introducing the
essential statement of the formal definition of a group:
\begin{itemize}
   \item A group is a non-empty set $G$ and an operation $+.$ The set $G$ is
closed under the operation $+.$ Regardless of which two elements $a, b$ from $G$
are taken, performing the operation on them gives an element in $G$ (i.e.
$a+b=c$, and $c$ lies in $G$).
   \item For all elements $a, b$ and $c$ in $G$: $(a+b)+c = a+(b+c)$.
   \item There exists an element $e$ in $G$ that behaves neutrally with respect
to the operation $+$. That means that for all a in the set $G: ~a+e = e+a = a.$
   \item For each element $a$ in $G$ there exists a so-called inverse element $-a$ 
($-a$ also lies in $G$) such that: $a+(-a) = (-a)+a = e$.
\end{itemize}
If also $a+b = b+a$ for all $a, b$ in $G$ then we call the group an {\em
Abelian} group. An operation denoted as $+$ indicates an {\em additive} group;
if the operation is denoted as $\cdot$, we speak of a {\em multiplicative}
group.

@@

An elliptic curve over ${\mathbb Z}_p$ is defined by an equation of the
following form: $$ y^2 \; ({\rm mod} \; p) = x^3 + ax + b ({\rm mod} \; p) $$
(thus: equality in the field ${\mathbb Z}_p$), where $a, b$ are in ${\mathbb
Z}_p$ and $4a^3 + 27b^2$ mod $p$ is not equal to zero. For fixed chosen numbers
$a$ and $b$ in ${\mathbb Z}_p$, this equation has the pair of solutions $$ {\bf
E} = \left\{(x,y) \left| \begin{array}{c} x {\rm ~and~} y {\rm ~are~in~}
{\mathbb Z}_p {\rm ~and~} \\ y^2  \equiv x^3 + ax + b \; ({\rm mod~} p) {\rm
~and~} \\ 4a^3 + 27b^2 \not\equiv 0 \; ({\rm mod~} p)\end{array} \right.
\right\}, $$ i.e. the set ${\bf E}$ consists of all pairs $x$ and $y$ that are a
solution (in ${\mathbb Z}_p$) of the above equation. It must be noted that the
numbers $a, b$ and $p$ determine which pairs $(x,y)$ lie in the set ${\bf E}$.
This means that $a, b$ and $p$ specify this set. The elements $(x,y)$ in ${\bf
E}$ are called points on the elliptic curve. In addition, ${\bf E}$ has one more
element $O$ (the so-called point in infinity). The set ${\bf E}$ is usually
called an elliptic curve. We can now define an operation (also denoted as $+$,
although it is not the standard/usual addition of real numbers) on two elements
in ${\bf E}$ such that the operation delivers an element that also lies in ${\bf
E}$. The set ${\bf E}$ is therefore closed under the operation $+$. We can show
that ${\bf E}$ is a group. The neutral element of the group ${\bf E}$ is the
point in infinity $O$. Thus, for every two points $(x_1,y_1)$ and $(x_2,y_2)$ on
the elliptic curve ${\bf E}$, there exists a point $(x_3,y_3)$ on ${\bf E}$ such
that the operation $+$ complies with the following: $(x_1,y_1) + (x_2,y_2) =
(x_3,y_3)$. Under certain circumstances, these points may also be equal to the
point in infinity. Thus, when we speak of a point $P$ on an elliptic curve ${\bf
E}$, we mean that $P = (x,y)$ and $(x,y)$ lies in the set ${\bf E}$. Any two
points on an elliptic curve specified by $a, b$ and $p$ can therefore be added
and the result is a point that also lies on the same elliptic curve. We must
note that ${\bf E}$ can have the following meanings:
\begin{itemize}
   \item the set ${\bf E}$ of solutions pairs $(x,y)$ for an equation including the point $O$
   \item the group ${\bf E}$ (with the operation ''addition of $(x_1,y_1)$ and $(x_2,y_2)$'')
   \item the elliptic curve ${\bf E}$ (which is actually the same as the group ${\bf E}$)
 \end{itemize}

@@

For cryptography, the important fact is that, for very large
numbers, it appears to be extremely difficult to use a given point
$Q$ on an elliptic curve to determine which two points have to be
added to obtain $Q$.

For large numbers $a, b$ and $p$ ($p$, for
example, has a length of more than $160$ bits), the computer can
easily add the point $P$  $m$ times after another, i.e. to
determine the point $P + P + \cdots + P = Q$ in an incredibly
short space of time (in a few fractions of a second) ($m$ summands
$P$). Rather than $P + P + \cdots + P = Q$ ($m$ summands $P$) we
also write $mP = Q$. If we have a point $P$ and a point $Q$, which
both lie on the same elliptic curve, no procedure is known that
enables us --- within an acceptable space of time --- to determine
the number $m$ (assuming it actually exists) for which $mP = Q$.
This is referred to as the ''elliptic curve discrete logarithm
problem'' (abbreviated to ECDLP).

We must note that not all elliptic curves are equally secure. This
means that we must choose the parameters $a$ and $b$ carefully
when defining a curve. For certain classes of elliptic curves, it
is possible to solve the ECDLP easier than in the general case.
Cryptographically unsuitable elliptic curves are called {\em abnormal}
curves (these are curves over ${\mathbb Z}_p$ for which the set
${\bf E}$ has precisely $p$ elements) and the {\em supersingular} curves
(the curves for which we can reduce the calculation of the ECDLP
to calculating the ''standard'' discrete logarithm in other finite
fields, i.e. simplify the calculation). There are therefore
cryptographically good and bad curves. However, for given
parameters $a$ and $b$ we can, with some difficulty, establish
whether or not the resulting elliptic curve is useful for
cryptographic purposes. The curves used in cryptography are
usually provided by experts. They ensure that the elliptic curves
they classify as secure satisfy the current security requirements.

For secure curves, the parameter $p$ determines how long it takes
to solve the ECLDP on this curve. The larger the parameter $p$,
the longer it takes to solve the problem. Experts recommend a bit
length of over $200$ bits for the parameter $p$. This makes it
clear why elliptic curves are so interesting for cryptography.
Because the parameter $p$ also determines the time required to
perform the signature/encryption procedure when using elliptic
curves in cryptography. The time taken to generate a pair of keys
also depends on $p$. Thus, small values (few bits) are desirable
here (in order to minimise the run times for the procedures);
however, the required security must still be maintained. For
example, with a length of $200$ bits for $p$, a {\em good}
elliptic curve is just as secure as an \index{RSA!module} RSA
module of over $1024$ bits in length (at least according to the
current state of research). The reason for this is that the
quickest algorithms for solving the {\em elliptic curve discrete
logarithm} problem have an exponential run time --- unlike the
subexponential run times that the best factorisation algorithms
currently have (number sieve, quadratic sieve or factorisation
with elliptic curves). Therefore, the parameters for cryptographic
procedures based on the problem of {\em factorising whole numbers}
must be greater than the parameters for cryptographic procedures
based on the ECDL problem.

@@

The {\em elliptic curve discrete logarithm problem} (ECLDP) forms the basis for
elliptic-curve cryptography. Various signature procedures are based on this.
What they have in common is how they use the public parameters and how these
parameters are used to generate secret and public keys:
\begin{itemize}
    \item The parameters of the elliptic curve E, i.e. a prime number $p$ that
determines over which field ${\mathbb Z}_p$ the elliptic curve ${\bf E}$ is
defined, as well as the two numbers $a$ and $b$ in ${\mathbb Z}_p$.
    \item A point $G=(x,y)$ that lies on the elliptic curve ${\bf E}$.
    \item A prime number $r < p$ for which $rG=O$ (i.e. $G$ added $r$ times
gives the neutral element of the group ${\bf E}$) and $r$ is a factor of $\#{\bf
E}$ (where $\# {\bf E}$ is the number of elements in the set ${\bf E}$). The
point $G$ therefore has the order $r$ and is the generator of a cyclic subgroup
of ${\bf E}$ with the order $r$.
    \item The number $k = \#{\bf E}/r$ ($k$ is called the {\em cofactor}).
\end{itemize}
The parameters $a, b, p, G, r$ and $k$ listed above are called
{\em domain parameters}.\index{Domain parameters} They determine on which elliptic curve ${\bf
E}$ and in which cyclic subgroup of ${\bf E}$ a signature
procedure has been ''used''.
